The host galaxy of the short GRB~111117A at $z = 2.211$\thanks{Based on observations collected at ESO/VLT under programme 088.A-0051 and 091.D-0904, at TNG under programme A24TAC\_38, at Gemini North under programme GN-2011B-Q-10 and GTC under programme GTC43-11B.}

### 2017-11-26 referee_reply.txt

Abstract: -"between 43 and 71 percent of all sGRB redshifts should be missed" --
what do you mean by all sGRB redshifts? all sGRBs that occur in the Universe,
all of those detected by Swift, or all based on the sample of short GRBs with
known hosts? I think it is the latter, but I would definitely clarify in the
abstract and when you get to this section in the text.

JS: We have added that this is for the Swift-detected bursts, both in the abstract and in the text.

Section 1:

--1st sentence: Define sGRB, NS.

JS: Done.

--For detection of kilonovae, also cite Berger+ 2013.

JS: Definitely. This was an oversight.

--"duration T_90 < 2" - make sure to qualify with 2 sec

JS: Done.

--Here and throughout, there is a tendency to use sGRB/short GRB and lGRB/long
GRB. I suggest using one or the other throughout the text.

JS: We have streamlined this a bit, but have kept the wording "long and short GRBs"

--"The bulk of these localizations have associated galaxies at relatively low redshifts... it is
arguably biased towards lower redshifts" -- This is not necessarily true, as the
low redshifts could be a natural consequence of relatively long delay times.

JS: We have cautioned the wording.

--"is tracing with host stellar mass rather than star-formation rate" --
Previous analysis of host galaxy properties have actually found that the short
GRB rate likely traces both stellar mass and star formation (see Berger 2014 and
references therein)

JS: We have changed the wording to reflect this.

--"will provide new a promising ways" -- delete "a"

JS: Done

Section 2: --The authors mention there were 4 separate epochs of spectroscopy -
is any spectral evolution apparent? This would at least be consistent with the
assumption that there is negligible afterglow contribution from epoch 1.

JS: We do not see any spectral evolution. This is now explicitly stated in the text.

--Since the spectra were taken across different nights and under different seeing
conditions, details of flux calibration for the combined spectrum should be
included.

JS: We have added a description of this.

--Is it surprising that [OIII] 4959 is not detected, given such a
strong detection of [OIII] 5007?

JS: [OIII] 4959 is detected, but very close to a sky-line. This has now been added to the text.

--"...because the slit is not aligned along the
long axis of the host, this value is an upper limit" - I believe this should be
a lower limit

JS: Correct! This has been changed and expanded.

--In calculating beta_OX, state the actual limits that were used
and in what band. Are these the deepest limits available?

JS: To our knowledge, these are the deepest limits available, yes. The limits have now been explicitly stated.

--"Both of these
authors report z-band photometry that is brighter by 0.8 mag to 1.0 mag compared
to our value" -- Are the authors comparing their Gemini z-band to z-band
measurements in the other papers? Or the VLT measurement to the measurements in
other papers? If the former, the previous values are actually brighter by
1.2-1.4 mag; if the latter, the values are discrepant by 0.7 to 0.9 mag. I could
not figure out the quoted values of 0.8 to 1.0 mag.

JS: We have changed this to ~ 1 mag. Both our z-band measurements are fainter than the z-band magnitudes reported by other authors.

Fig. 3: best-fist --> best-fit

JS: Done

Section 3: --"The sub-samples comrprise of 333 long, 19 short..." - the papers
where the data came from should be cited. 
These references were in the caption, but have now been repeated in the text. 
--"had a luminosity of 44.95" --> should appear as a power of 10 and with units

JS: This has been updated.

--"GRB 111117A has a 96% probability of being a sGRB"

JS: Done

-- The terms sGRB and non-collapsar seem to be used interchangeably here. From the burst properties it certainly
seems like a sGRB. To be more accurate I would say "non-collapsar"

JS: We have removed the usage of non-collapsar and stuck with sGRB.

--"... it occurred when the universe was younger by 1.8 Gyr compared to any
other non-collapsar GRB ever detected"

JS: Done

-- again, how do you find non-collapsar? There is no way to know for certain and
the Bromberg+ 2013 method has been proven unreliable from host galaxy
properties. If you do want to rely on Bromberg+ 2013, what percent probability
do you consider as a "non-collapsar GRB"?

JS: We have stated a f_NC > 50 per cent requirement of the Bromberg classification to consider a given burst short.

--"This number is 3.2 Gyr for the next-highest spectroscopic redshift" - This
sentence needs a citation and would be helpful to state what burst this is

JS: We have expanded this entire section a bit to be more complete.

--Even though the offset does not change much, it is worth recalculating in this
paper for completeness.

JS: Done

--"Along with the absence of dust" -- This is not yet proven in this paper. I
suggest revising to "If the assumption of no dust is correct..."

JS: This has been changed to "low dust content"

--The scenario in which the density of material is high to explain the high NH
would imply a bright afterglow, although would appear faint at z=2.211. Are the
afterglow limits still consistent with a high-density environment at z=2.211?

JS: The X-ray afterglow is not strongly dependent of the density and because of the optical
darkness, dust extinction locally to the GRB can accomodate a large brightness
interval. We have mentioned this in text.

@@

Fig. 4: A comparison of the light curve for GRB 111117A to the identical figure
in Margutti+ (their Fig 6) looks like the luminosity does not change, even
though it should be significantly higher given the revised redshift. In fact,
the Margutti+ light curve exceeds 10^48 erg s^-1 near the start, while the light
curve in Fig. 4 does not. Is the luminosity calculated correctly, or are they in
different energy bands? -Does this include the Chandra data as well? If so, it
would be good to include that in the caption.

JS: Margutti+ didn't use the light curve from the Swift Burst Analyser. Her early light
curve is ~0.18 dex brighter. The shape is also a bit different. k-correction:
Margutti+ fixed the photon index to 2.2. The Burst Analyser provides estimates of
the photon index for each observation. At the time of the first observation, the
Burst Analyser gives a photon index of 0.93. The k-correction for a power-law
shaped spectrum is: (1+z)^(Gamma-2). The difference in these parameters
corresponds to a difference in the k-correction of 0.63 dex. The different
distances gives a luminosity difference of 0.58 dex (differences in the
cosmology are taken into account).
All in all, logL(us) - logL(Margutti) = (0.58 - 0.18 - 0.63) dex = -0.23 dex.
Comparing her plot with ours gives difference of ~-0.2 dex, which is identical
to the expected difference.

Fig. 5: What are the lines and what is the red curve that appear on the top and
side of the plot?

JS: The caption has been expanded to better explain the different
components.

Section 5: -"this places a hard upper limit on the initial
separation..." - Is there a citation for the equation used to calculate this?

JS: The equation is in the Postnov reference. The actual calculation is included in
the online repository, with reference to the exact equation.

-This may be outside the scope of this work, but how does
the initial separation change under different mass assumptions (unequal mass
binary, or more/less massive binary)?

JS: We have included the calculation for different mass assumptions and changed the text accordingly.

-The authors should note that this
discussion assumes that GRB 111117A did not form dynamically in a globular
cluster as suggested by some works (e.g., Lee+ 2010; Church+ 2011)

JS: We do not assume anything about the formation channel, only that the neutron
star binary was already in place at the beginning of the universe. We have for
completeness mentioned the possible dynamical origin of neutron star binaries.

Section 6: -"The high NH is difficult to reconcile with the large projected host
offset" - How large is the offset at z=2.211? -"for between 43 and 71 percent of
sGRB hosts" -- should be revised to "sGRBs with known or well-characterized
hosts"

JS: We have changed the wording.
-------------------

### 2017-11-26 GRB111117A.tex
Last edits

The large offset of GRB\,111117A relative to the host center derived in
\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
galaxy-scale gas being the source of the X-ray absorption. Along with the low
dust content of the host, the large offset from the host center indicates that
the high $N_\mathrm{H,X}$ arises because the density in the GRB surroundings is
high (or possibly because the light from the afterglow transverses localized
regions of dense gas), see for example,
\citep[e.g.][]{Galama1998,Watson2013,Krongold2013}. Alternatively, it has been
hypothesised that a significant contribution to the observed X-ray
$N_\mathrm{H,X}$ could come from the diffuse intergalactic medium and the
intervening systems along the line of sight of the GRB \citep{Campana2012,
	Arcodia2016}, although see \citep{Watson2013, Krongold2013} for counter
arguments.

Even assuming a low dust-to-metals ratio as typically observed in long GRB
afterglow sightlines \citep{Galama2001, Schady2010, Covino2013}, the
$N_\mathrm{H,X}$ value derived from the X-ray spectrum corresponds to
significant extinction along the afterglow line of sight ($A_V \gtrsim 1$~mag),
which is contrasted to the absence of dust found from the SED fit and supported
by the detection of \lya{}. Such a discrepancy between the extinction derived
from the GRB afterglow and the one obtained using galaxy-wide measures has also
been observed occasionally for lGRBs \citep{Perley2013a}. For the one sGRB where
both parameters were measured (GRB~130603B; \citealt{DeUgartePostigo2014d}),
they were found to be consistent with $A_V \sim 1$~mag.

@@

The redshift distribution of GRBs is a powerful informant on both the conditions
which drive the formation of these events, but also on the potential influence
these cosmic explosions have on the evolution of the universe. Due to the
elevated brightness of lGRBs compared to sGRBs \citep{Berger2014} and their
tendency to be associated with the star-forming and therefore dense regions in
their hosts \citep{Fruchter2006, Lyman2017}, the redshifts of lGRBs are easier
to measure than for sGRB counterparts, where only a single burst has a redshift
measurement from the GRB afterglow \citep{Cucchiara2013, DeUgartePostigo2014c}.
Correspondingly, the redshift distribution of sGRBs is still substantially
unconstrained compared to that of lGRBs \citep[e.g., see][]{Jakobsson2012a,
	DAvanzo2015, Perley2016d}.

>>>>

The large offset of GRB\,111117A relative to the host center derived in
\citet{Margutti2012} and \citet{Sakamoto2013} is difficult to reconcile with
galaxy-scale gas being the source of the X-ray absorption. Along with the low
dust content of the host, the large offset from the host center indicates that
the high $N_\mathrm{H,X}$ arises because the density in the GRB surroundings is
high (or possibly because the light from the afterglow transverses localized
regions of dense gas), see for example,
\citep[e.g.][]{Watson2013, Krongold2013}. Alternatively, it has been
hypothesised that a significant contribution to the observed X-ray
$N_\mathrm{H,X}$ could come from the diffuse intergalactic medium and the
intervening systems along the line of sight of the GRB \citep{Campana2012,
	Arcodia2016}, although see \citep{Watson2013, Krongold2013}.

Even assuming a low dust-to-metals ratio as typically observed in long GRB
afterglow sightlines \citep{Galama2001, Schady2010, Covino2013}, the
$N_\mathrm{H,X}$ value derived from the X-ray spectrum corresponds to
significant extinction along the afterglow line of sight ($A_V \gtrsim 1$~mag),
which is contrasted to the absence of dust found from the SED fit and supported
by the detection of \lya{}. Such a discrepancy between the extinction derived
from the GRB afterglow and the one obtained using galaxy-wide measures has also
been observed occasionally for lGRBs \citep{Perley2013a}. For the one sGRB where
both parameters were measured (GRB~130603B; \citealt{DeUgartePostigo2014b}),
they were found to be consistent with $A_V \sim 1$~mag.

@@

The redshift distribution of GRBs is a powerful informant on both the conditions
which drive the formation of these events, but also on the potential influence
these cosmic explosions have on the evolution of the universe. Due to the
elevated brightness of lGRBs compared to sGRBs \citep{Berger2014} and their
tendency to be associated with the star-forming and therefore dense regions in
their hosts \citep{Fruchter2006, Lyman2017}, the redshifts of lGRBs are easier
to measure than for sGRB counterparts, where only a single burst has a redshift
measurement from the GRB afterglow \citep{Cucchiara2013, DeUgartePostigo2014b}.
Correspondingly, the redshift distribution of sGRBs is still substantially
unconstrained compared to that of lGRBs \citep[e.g., see][]{Jakobsson2012a,
	DAvanzo2015, Perley2016d}.
